### 2022-01-31 include/abstract.tex
draft onset added to manuscript

\chapter*{Abstract}
%\phantom{Abstract}

The population of corals throughout the world has declined dramatically under the effect of rising water temperature and many coral species are now listed as threatened. In addition to ocean warming, other threats to corals include (but are not limited to) disease outbreaks, intensifying hurricanes and sediment disturbances. In particular, Florida's Coral Reef (FCR) is currently facing a multi-year outbreak of the stony coral tissue loss disease (SCTLD). First observed in 2014 during the monitoring of dredging operations in the Port of Miami, the disease has now spread through the entirety of FCR and has been observed in numerous territories of the Caribbean. Furthermore, hurricanes tend to strike Florida during the reproduction period of the corals, when thousands of larvae are released in the water column. In addition to causing physical harm to corals, hurricanes significantly modify ocean currents and might therefore disturb larval exchanges between reefs. A high resolution ocean model, SLIM, is used to build a biophysical model to study the aforementioned threats to FCR. First, a coupled epidemio-hydrodynamic model is developed to study both the onset and the propagation mechanisms of the SCTLD. Furthermore, a coupled wave-current model is devised to study the impact of Hurricane Irma (2017) on transport processes in the Florida Keys. This study brings new insight on the impact of hydrodynamics on the propagation of SCTLD as well as the impact of hurricane on coral ecosystem.

%The population of corals throughout the world has declined dramatically and many coral species are now listed as threatened Avoiding global warming being no longer an option, reef managers now focus on damage control and restoration strategies. The Florida Reef Tract, currently facing an unprecedented coral disease outbreak and where healthy corals grown in nurseries are replanted on damaged reefs, is a good application case of such reef management strategies. In order to maximize the efficiency of these efforts, a good knowledge of exchanges of infectious materials and larvae between reefs is required. However, connectivity patterns between reefs remain mostly unknown. To address this issue, we will use the high-resolution ocean model SLIM to simulate connectivity patterns between reefs and then identify reefs with the highest restoration potential as well the ones that are most at risk of being infected by the outbreak.

%The dispersal of coral larvae away from their natal habitats is an important process for coral reef ecosystems, but remains poorly understood and hard to gauge. This thesis presents a spatially explicit numerical modelling approach to simulate larval dispersal in Australia's Great Barrier Reef (GBR), and in doing so to estimate connectivity between the thousands of reefs that comprise it. A high resolution, finite element ocean model, SLIM, is used to build a biophysical model of larval dispersal capable of estimating larval exchange between reefs in large sections of the the GBR. Tools from network science are employed to investigate the community structure of the GBR's connectivity network, and these tools are used to reveal the presence of clusters of highly inter-connected reefs. Subsequently, the model is used to estimate the extent of connectivity between reefs of different morphologies and depths in the GBR. Finally, the possible impacts of climate change on connectivity in the GBR are assessed, and estimates are made of the likely changes to coral larval dispersal patterns over the coming century.

%Titan, Saturn’s largest moon, is a unique object in the Solar system as much as it has a substantial atmosphere, a surface with a complex interplay of geological processes and an outer ice shell overlyinga subsurface ocean. The climate on this icy satellite boasts a multi-phase hydrological cycle where methane plays a role similar to that of water on Earth. Surface lakes and seas filled with liquid methane and ethane are found in the polar regions. The subsurface ocean, on the other hand, is filled with liquid water. This liquid layer lying beneath Titan’s surface allows for the large surface deformation observed.

%The first objective of this thesis was to adapt an Earth-based geophysical and environmental model, SLIM (www.climate.be/slim), to Titan’s specific conditions in order to study the tidal motion in the surface lakes and seas. The modified model was applied to the largest lake in the southern hemisphere, Ontario Lacus, and the two largest seas, Kraken and Ligeia Maria, which are located in the northern hemisphere. The predicted surface elevation and velocity fields are part of the data needed to develop an exploration mission focusing on the surface lakes and seas of Titan. The normal modes of Ontario Lacus were also numerically studied. While resonantly forced modes could generate significant liquid motion, the natural periods are much shorter than the period of the astronomical forcings. Therefore, only atmospheric forcings could resonantly force the normal modes. Strong wind conditions corresponding to a stormy event are required to generate a surface elevation resulting in significant shoreline variations.

%Then, the model was used to study the tides of Titan’s global subsurface ocean. To this end, the shallow water equations and the 3D hydrostatic equations under the Boussinesq approximation were modified in order to take into account the ice shell lying at the top of the ocean. The method was adapted to solve the equations on a sphere. The effects of the shell are represented by adding a surface friction and a surface pressure term. The shell decreases the ocean surface elevation and slows down the flow without modifying significantly the global patterns of these fields. The magnitude of these variations depends on the mechanical behaviour of the ice shell. Surface and bottom heat fluxes play a significant role in the liquid motion of the ocean. The interactions between the tidal motion and the thermally driven flow resulting from the surface heat flux were studied by means of the 3D version of SLIM. The surface heat flux significantly impacts the velocity field, both in terms of magnitude and orientation while the influence on the ocean’s surface elevation is small. These results could be useful for astrobiologists paying attention to the ocean habitability and to validate hypotheses about Titan’s internal structure.