### 2022-02-07 thesis.tex

\documentclass[twoside,12pt,a4paper,openright]{memoir}
\maxsecnumdepth{subsubsection}			%Automatically number sections down to ``section'' level
\setsecnumdepth{subsubsection}
% \hyphenation{de-leer-snij-der}
\usepackage[utf8]{inputenc}			%Optimal font encoding for most latin languages
\usepackage[T1]{fontenc}
\usepackage{changepage}
\usepackage{pgfplots}
\usepackage{bm}
\usepackage{lipsum}
\input{include/specific_headers.tex}
\newcommand{\GG}[1]{}
\pgfplotsset{compat=1.15}
\usepackage[pdftex,linktocpage]{hyperref}		%For hyperlinks within doc, and setting doc info
\hypersetup{pdftitle={How modeling tools can inform coral reef management in Florida}}
\hypersetup{pdfauthor={Thomas DOBBELAERE}}
\hypersetup{pdfsubject={How modeling tools can inform coral reef management in Florida}}
\hypersetup{pdfkeywords={Biophysical modeling, Florida Coral Reef, SCTLD, SLIM}}
% % Comment out next hypersetup section for print version:
%\hypersetup{
%  colorlinks=true,
%  citecolor=blue,
%  linkcolor=blue,
%  urlcolor=blue
%}
\hypersetup{
  colorlinks=true,
  citecolor=black,
  linkcolor=black,
  urlcolor=black
}
% % % %   citebordercolor={0.8 0.8 0.8},
% % % %   filebordercolor={0.8 0.8 0.8},
% % % %   linkbordercolor={0.8 0.8 0.8},
% % % %   menubordercolor={0.8 0.8 0.8},
% % % %   urlbordercolor={0.8 0.8 0.8},
% % % %   runbordercolor={0.8 0.8 0.8}

% \usepackage{cmbright}				%TODO: this doesn't compile when I add it *** Font: Computer modern bright
% \usepackage{sfmath}				%Display all maths with sans serif font

\begin{document}
\frontmatter
  \input{include/title}
  \clearemptydoublepage
  \input{include/abstract}
  %\clearemptydoublepage
  \input{include/thanks}
  \clearemptydoublepage
  \tableofcontents
  \clearemptydoublepage
  % \input{include/glossary}
  % \clearemptydoublepage
  \mainmatter 
%%%%
\mychapter{Introduction} \label{chap:intro}
 \input{chapters/intro/intro}    
%%
%%%%%% 
\mychapter{Modeling the spread of SCTLD in Florida} \label{chap:sctld}
\chaptermark{Modeling the spread of SCTLD in Florida}  
\input{chapters/sctld/sctld}
\mychapter{Transmission of SCTLD from the Marquesas to the Dry Tortugas}\label{chap:drto}
\chaptermark{Transmission of SCTLD from the Marquesas to the DRTO}  
\input{chapters/drto/drto}
\mychapter{The onset of the SCTLD outbreak and the PortMiami Deep Dredge Project} \label{chap:onset}
\chaptermark{Onset of SCTLD and PortMiami Deep Dredge Project}  
\input{chapters/onset/onset}
\mychapter{Impacts of hurricanes on wave-induced ocean transport processes} \label{chap:irma}
\chaptermark{Wave-induced transport under hurricanes}  
\input{chapters/irma/irma}
%%%%%%%%%%%
%
%%%%%  
%%%
  \mychapter{Conclusions and perspectives} \label{chap:conclusion}
  \input{chapters/conclusions/conclusions}    
%%
%
  \clearemptydoublepage
  \bibliography{include/bibliography.bib}

\end{document}