### 2022-02-13 chapters/onset/onset.tex

This chapter is based on the following article:
\begin{list}{}{%
\setlength{\topsep}{0pt}%
\setlength{\leftmargin}{0.23in}%
\setlength{\listparindent}{-0.23in}%
\setlength{\itemindent}{-0.23in}%
\setlength{\parsep}{\parskip}%
}%

\item \textbf{Dobbelaere, T.}, other co-authors TBD, \& Hanert, E. Early transmission dynamics of a deadly coral disease and its connection with the PortMiami Deep Dredge Project. (planned submission to \textit{Marine Pollution Bulletin})
\end{list}

\begin{abstract}
    For the last 8 years, Florida's Coral Reef (FCR) has suffered from severe coral loss caused by stony coral tissue loss disease (SCTLD). The outbreak reportedly initiated near Virginia Key in September 2014, during the deepening of the Port of Miami (PoM) shipping channel, that took place between November 20, 2013 and March 16, 2015. The disease then spread to the entirety of FCR, likely through waterborne transmission. Although the causative agent of SCTLD remains unknown, there is evidence that sediment can act as vector for the disease. Here we evaluate whether sediments produced or resuspended during dredging operations could have reached reefs where the disease was first reported and assess whether the timing of these sediment fluxes was consistent with the observed spread of the disease. We evaluated the transport of sediments caused by the expansion of PoM between November 2013 and September 2014 using a quasi-3D sediment transport model forced by currents from the high-resolution coastal ocean model SLIM. A particular attention was paid to the non conventional dredging operations performed without pumping the produced chopped rock particles. Our result suggest that the dredging had close to no direct impact on the coral reefs of Virginia Key. However, a significant fraction of the sediments produced reached monitoring sites where the disease was reported prior to September 2014, suggesting that the dredging might have played a role in the onset of the outbreak at these sites. Furthermore, using a biophysical disease agent transport model, we show that waterborne transmission from one of these sites to Virginia Key was possible before September 2014. This study brings new insight on the role that the deepening of PoM might have played in the onset of one of the worst coral outbreaks on record in the Caribbean.
\end{abstract}

%\newpage

\input{chapters/onset/article}